European or American? A re-examination of the transatlantic influence over the IMF

### 2025-10-04 manuscript/manuscript.qmd
Update packages and disable DESCRIPTION (#84)
* Update packages

* Update tinytable outputs

* Remove DESCRIPTION

* Downgrade Matrix (later version not usable)

* Add preamble for new tinytable

* update scree plot

* README: use test-manuscript as file example

* test-manuscript: use local edition 3

* Update snapshots

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
date-format: |
    | DD MMMM YYYY
authors:
  - name: Dianyi Yang
    email: dianyi.yang@politics.ox.ac.uk
    orcid: 0009-0004-4652-3429
    corresponding: true
    affiliations:
      - ref: ox    
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
    affiliations:
      - ref: bw
    equal-contributor: true
  - name: Zhenyi Chen
    orcid: 0009-0008-2910-5187
    affiliations:
      - ref: rmu
    equal-contributor: true
affiliations:
  - id: ox
    name: Department of Politics and International Relations, University of Oxford
  - id: bw
    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
  - id: rmu
    name: School of International Studies, Renmin University of China
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
          \AtAppendix{\counterwithin{table}{section}}
abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance. While the Fund is widely regarded as dominated by “Western” powers, most existing research either focuses exclusively on American influence or treats the West as a unified bloc, masking important variation between its two key actors. Drawing on panel data from 1980 to 2010, we construct separate measures of US and European influence using principal component analysis of commonly used alignment indicators—UN voting patterns, bilateral trade, and banking exposure—and estimate a series of Tobit and Probit models to assess their impact on four key IMF lending outcomes: loan-to-GDP ratios, participation rates, loan approvals, and conditionality. We find that European influence is strongly associated with higher loan volumes, more frequent participation, and greater likelihood of loan approval, while US influence is not. However, American ties are significantly correlated with fewer conditions attached to loans. Additional analysis shows that European influence is not driven by regional or colonial favouritism: neither European nor African borrowers benefit more from European ties—and may, in fact, fare worse—than other countries with comparable alignment. These findings complicate existing assumptions about IMF decision-making, highlighting the need to distinguish between types of Western influence. They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality. In doing so, the paper contributes to debates on hegemony, institutional power, and intra-Western dynamics in global economic governance."
keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
execute:
  echo: false
  warning: false
  error: false
  message: false
---

>>>>

---
title: European or American? A re-examination of the transatlantic influence over the IMF
date: last-modified
date-format: |
    | DD MMMM YYYY
authors:
  - name: Dianyi Yang
    email: dianyi.yang@politics.ox.ac.uk
    orcid: 0009-0004-4652-3429
    corresponding: true
    affiliations:
      - ref: ox    
  - name: Tong Hu
    orcid: 0009-0003-5234-8164
    affiliations:
      - ref: bw
    equal-contributor: true
  - name: Zhenyi Chen
    orcid: 0009-0008-2910-5187
    affiliations:
      - ref: rmu
    equal-contributor: true
affiliations:
  - id: ox
    name: Department of Politics and International Relations, University of Oxford
  - id: bw
    name: School of International Relations and Diplomacy, Beijing Foreign Studies University
  - id: rmu
    name: School of International Studies, Renmin University of China
format:
  #native: default
  #docx: default
  orcid-pdf:
    keep-tex: true
    number-sections: true
    include-in-header:
      - text: |
          \newcommand{\centerstart}{\begin{center}}
          \newcommand{\centerend}{\end{center}}
          \usepackage{amsthm, thm-restate}
          \usepackage{}
          \newtheorem{hyp}{Hypothesis}
          \newtheorem{subhyp}{Hypothesis}[hyp]
          \renewcommand{\thesubhyp}{\thehyp\alph{subhyp}}
          \usepackage{chngcntr}
          \usepackage{apptools}
          \AtAppendix{\counterwithin{figure}{section}}
          \AtAppendix{\counterwithin{table}{section}}
      - file: extra/preamble.tex
abstract: "This paper investigates whether the United States or Europe holds greater influence over the International Monetary Fund (IMF)—a longstanding yet underexamined question in the study of global economic governance. While the Fund is widely regarded as dominated by “Western” powers, most existing research either focuses exclusively on American influence or treats the West as a unified bloc, masking important variation between its two key actors. Drawing on panel data from 1980 to 2010, we construct separate measures of US and European influence using principal component analysis of commonly used alignment indicators—UN voting patterns, bilateral trade, and banking exposure—and estimate a series of Tobit and Probit models to assess their impact on four key IMF lending outcomes: loan-to-GDP ratios, participation rates, loan approvals, and conditionality. We find that European influence is strongly associated with higher loan volumes, more frequent participation, and greater likelihood of loan approval, while US influence is not. However, American ties are significantly correlated with fewer conditions attached to loans. Additional analysis shows that European influence is not driven by regional or colonial favouritism: neither European nor African borrowers benefit more from European ties—and may, in fact, fare worse—than other countries with comparable alignment. These findings complicate existing assumptions about IMF decision-making, highlighting the need to distinguish between types of Western influence. They also suggest that Europe wields broader institutional influence over access to IMF resources, while US power manifests more selectively through conditionality. In doing so, the paper contributes to debates on hegemony, institutional power, and intra-Western dynamics in global economic governance."
keywords: International Monetary Fund, hegemony, global economic governance, US influence, European influence, conditionality
execute:
  echo: false
  warning: false
  error: false
  message: false
---
